---
title: "A Core Isotope Workflow with ggbothbar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Core Isotope Workflow with ggbothbar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

`ggbothbar` supports a compact workflow for replicate isotope measurements:
calculate enrichment relative to a reference group, summarise uncertainty in
both dimensions, and label group means.

## Prepare replicate measurements

```{r data}
library(ggbothbar)

isotopes <- data.frame(
  type = rep(c("reference", "orchid A", "orchid B"), each = 3),
  d13C = c(-27.1, -26.8, -27.0, -23.5, -23.1, -23.3, -20.9, -21.2, -21.0),
  d15N = c(2.0, 2.3, 2.1, 5.8, 6.1, 5.9, 8.2, 7.9, 8.1)
)
```

`calc_enrichment()` subtracts each reference mean from every observation. The
default output names replace the leading `d` with `e`.

```{r enrichment}
enriched <- calc_enrichment(
  isotopes,
  delta = c("d13C", "d15N"),
  reference = "reference"
)

enriched
```

Existing output columns are protected by default. Use `overwrite = TRUE` only
when replacing them is intentional.

## Plot grouped uncertainty

Map `group` explicitly so each biological group is summarised separately.
`geom_errorbarb()` calculates the group mean and either standard deviation or
standard error along both axes. Its cap width is a physical size in millimetres,
so it remains visually stable when the plotting viewport changes.

```{r plot, fig.alt="Carbon and nitrogen enrichment means with two-axis standard errors"}
ggplot(
  enriched,
  aes(e13C, e15N, colour = type, group = type, label = type)
) +
  geom_point(alpha = 0.55) +
  geom_errorbarb(fun.errorbar = "se", errorbar_tip_size = 2) +
  stat_mean_point(size = 2.5) +
  stat_mean_label(
    geom = "text",
    position = position_nudge(x = 0.25, y = 0.2),
    show.legend = FALSE
  ) +
  labs(
    x = label_isotope(13, "C", notation = "epsilon"),
    y = label_isotope(15, "N", notation = "epsilon")
  ) +
  theme_aca(base_size = 13, base_family = "sans")
```

For already summarised data, use `geom_errorbarb(stat = "identity")` and map
`xmin`, `xmax`, `ymin`, and `ymax` explicitly.
